---
title: "Webscraping Project - Analysising Metacritic Videogame Reviews"
author: "Jeremy Osir"
date: "10/4/2021"
output: html_document
---

#### Next steps: Clean up Chunk Outputs; Knit; Presentation!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Data Preprocessing ####

### 1.1 Load libraries
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
library(dplyr)
library(ggplot2)
library(readr)
library(skimr)
library(data.table)
library(formattable)
```

### 1.2 Load dataset
```{r load_data, cache=TRUE, warning=FALSE, message=FALSE}

game.reviews <- read.csv("./metacritic_VGreviews.csv") 

```

### 1.3 Clean data

```{r preprocessing}
# Review summary information about the game reviews dataframe
skim(game.reviews) 
  # 17K observations with 20 variables/features
  # there are no missing values in the 11 character columns
  # numeric columns have missing values, but those seem to be bc reviews are stored in different fields


#### Pre-Processing Steps ####
# 1. Convert "release date" column from character to date
game.reviews.df <- game.reviews %>% 
  mutate(., release_date = mdy(release_date)) # Dates in format: "Month - Day - Year"

sapply(game.reviews.df, class) # check variable classes


# 2. Create new column that extracts the year from release_date
game.reviews.df <- game.reviews.df %>% 
  mutate(., release_year = year(release_date))
  
# 3. The critic and user review scores are stored in different columns based on whether they are positive, mixed, or negative. Create ONE column to store all the critic reviews
game.reviews.df <-game.reviews.df %>% 
  mutate(., critic_score = case_when(
    is.na(critic_score_pos) & is.na(critic_score_mix) ~ critic_score_neg,
    is.na(critic_score_neg) & is.na(critic_score_pos) ~ critic_score_mix,
    is.na(critic_score_mix) & is.na(critic_score_neg) ~ critic_score_pos,
    TRUE ~ critic_score_pos
  ))

  # Two games do not have any critic reviews (Kalypso and Eggtooth Team). Exclude them from the dataset
game.reviews.df <-  game.reviews.df %>% 
  filter(., !is.na(critic_score))

# Next, repeat the same process for the user review columns
game.reviews.df <- game.reviews.df %>% 
  mutate(., user_score = case_when(
    is.na(user_score_pos) & is.na(user_score_mix) ~ user_score_neg,
    is.na(user_score_neg) & is.na(user_score_pos) ~ user_score_mix,
    is.na(user_score_mix) & is.na(user_score_neg) ~ user_score_pos,
    TRUE ~ user_score_pos
  ))

# The critic scores are given on a 1-100 scale, while the user scores are 1-10 (to 1 decimal place)
# Scale the critic-score by 0.1 so that it matches the user score format
game.reviews.df <- game.reviews.df %>%
  mutate(critic_score_scaled = critic_score/10)

## Genres are all stored in one column. Let's split them into seperate columns

  # Check the maximum no. of genres that a single game in the dataset has
max(lengths(strsplit(game.reviews.df$genres, ','))) # 12 genres

# Use 'str_split_fixed' function to split each genre in the genres column into individual columns
  # NOTE: since there are duplicate genres, let's plan to remove dupes after making the df longer
genre.columns <- str_split_fixed(game.reviews.df$genres, ",", 12)

# combine the genre
reviews.genres <- cbind(game.reviews.df, genre.columns)

# gather the dataframe and make it long
genres.long <- pivot_longer(reviews.genres, 
                            cols = "1":"12", names_to = "genre.ID", values_to = "genre.value")

# Each title's genres are captured in a separate row. 
# Remove any rows that have an empty genre.value and/or a duplicated value
genres.long <- genres.long %>% 
  filter(., !genre.value == "")


reviews.genres.long <- genres.long %>% 
  distinct(., title, platform, genre.value, .keep_all = TRUE)


```

## 2. Data Visualization ####
### Questions we want to investigate in this dataset

- How do user scores compare to critic scores - overall? Is there a correlation?
- How have user & critic reviews trended over time? 
- Which platforms have the highest reviews, on average according to 1) critics, 2) users?
- Which platform has the most games ranked in the top 100? Bottom 100?
- Which videogame genres scored highest in the eyes of 1) critics, 2) users?
- Which videogames are in the Hall of Fame? Which are in the the Hall of Shame?
- Do my favorite games perform better or worse in than the rest of the games?


### 2.1 Inspection: Compare distribution of user scores vs critic scores? Is there a correlation?
``` {r }
# Create some charts to explore the distribution  data
  # User score frequency plot
user.reviews.distr <-game.reviews.df %>% 
  ggplot(data = ., aes(x = user_score)) +
  geom_bar(position = "dodge", fill = "blue") +
  labs(title = "Distribution of Users' Videogame Review Scores",
       x= "User Review Score", y = "Frequency")+
  theme(plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),
        legend.position = "none")

user.reviews.distr

  # Critic score frequency plot
critic.reviews.distr <- game.reviews.df %>% 
  ggplot(., aes(x = critic_score_scaled, fill = "red")) +
  geom_bar(position = "dodge")+
  labs(title = "Distribution of Critics' Videogame Review Scores",
       x= "Critic Review Score", y = "Frequency") +
  theme(plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),
        legend.position = "none")
critic.reviews.distr

  # Overlap critic and user score distribution on one chart

# Let's observe density plots for the user and critic reviews
game.reviews.distr <- game.reviews.df %>% 
  pivot_longer(., cols = c(user_score, critic_score_scaled), names_to = "review_type", values_to = "review_score") %>% 
  ggplot(., aes(x = review_score, fill = review_type))+
  geom_density(alpha = 0.6) +
  labs(title = "Distribution of Videogame Review Scores\n Users vs Critics",
       x = "Review Score",
       fill = "Review Type")+
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))+
  scale_fill_discrete(labels = c("Critic Score", "User Score"))
game.reviews.distr

# Correlation between user and critic scores ####
scatter.user.critic.reviews <- game.reviews.df %>% 
  ggplot(., aes(x = critic_score_scaled, y = user_score))+
  labs(title = "Video Game Reviews: Critic Scores vs User Scores", x= "Critic Review Score", y = "User Review Score") +
  geom_point(size = 0.6, color = "red")+
  geom_smooth(method = "lm") +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))
scatter.user.critic.reviews




# Check correlation between critic score and user score
cor.test(game.reviews.df$critic_score_scaled, game.reviews.df$user_score, method = "pearson", conf.level = 0.95)
  # Pearson's correlation test indicates user and critic scores have a correlation = 0.53, and is statistically significant
  # Takeaway: User and Critic Scores are moderately positively correlated, with scores generally following the same pattern




```

### 2.2 Inspection: Examine how the average review scores given by users and critics has changed over time.
  Check whether the difference in review scores has got larger or smaller.
```{r}

# Title Reviews by Year - comparing critic scores vs user scores ####
users.critics.annual <- game.reviews.df %>% 
  group_by(.,release_year) %>% 
  filter(., !release_year == 2021) %>% 
  summarise(., avg.annual.critic.score = mean(critic_score_scaled),
            avg.annual.user.score = mean(user_score)) %>% 
  mutate(., score.difference = abs(avg.annual.critic.score - avg.annual.user.score))

user.critics.annual.chart <- users.critics.annual %>%
  pivot_longer(., cols = !c(release_year,score.difference), names_to = 'review_type', values_to = "review_score") %>% 
  ggplot(., aes(x = release_year, y = review_score, color = review_type))+
  geom_line() +
  labs(title = "Metacritic Videogame Review Scores by Release Year\n Users vs Critics: 1995 - 2020",
       x= "Release Year", y = "Review Score",
       color = "Review Type") +
  scale_color_hue(labels = c("Average Critic Score", "Average User Score")) +
  theme(axis.text.x = element_text(size = 10), axis.title.x = element_text(size = 12),
        axis.text.y = element_text(size = 10), axis.title.y = element_text(size = 12),
        plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))

user.critics.annual.chart

# Let's look at how much the scores have diverged over time
user.critics.score.diff <- users.critics.annual %>% 
  ggplot(., aes(x = release_year, y = score.difference))+
  geom_line() +
  labs(title = "Difference between average critic and user scores each year\n 1995-2020",
       x= "Release Year", y = "Difference") +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))

user.critics.score.diff



# Title Reviews by Decade ####
title.reviews.decade <- game.reviews.df %>% 
  filter(., !release_year == 2021) %>% 
  mutate(., release_decade = case_when(
    release_year %in% c(1990,1991,1992,1993,1994,1995,1996,1997,1998,1999) ~ "1990s",
    release_year %in% c(2000,2001,2002,2003,2004,2005,2006,2007,2008,2009) ~ "2000s",
    release_year %in% c(2010,2011,2012,2013,2014,2015,2016,2017,2018,2019) ~ "2010s",
    TRUE ~ "2020s")) %>% 
  group_by(., release_decade) %>% 
  summarise(., avg.annual.critic.score = mean(critic_score_scaled),
            avg.annual.user.score = mean(user_score)) %>% 
  pivot_longer(., cols = !release_decade, names_to = 'review_type', values_to = "review_score") 


title.reviews.decade.chart <- title.reviews.decade %>%
  ggplot(., aes(x = release_decade, y = review_score, fill= review_type, group = review_type))+
  geom_col(position = "dodge") +
  labs(title = "Metacritic Videogame Review Scores by Release Decade\n Users vs Critics",
       x= "Decade", y = "Average Review Score",
       color = "Review Type") +
  scale_color_hue(labels = c("Average Critic Score", "Average User Score")) +
  theme(axis.text.x = element_text(size = 10), axis.title.x = element_text(size = 12),
        axis.text.y = element_text(size = 10), axis.title.y = element_text(size = 12),
        plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))

title.reviews.decade.chart

# On average, user review scores have gone down over time.
# The 90s have the highest reviews by users and critics. (Partially due to fewer games being released)
# The average review scores dropped by almost 1- 1.5pts in the 2000s.
# By the 2010s, we start seeing more divergence between critics and users, with users being more critical.




```

## 2.3. Inspection: How has the number of user & critics providing reviews changed over time? 

```{r}
# How have number of user provided reviews trended over time? 
user.reviews.freq <- game.reviews.df %>%
  group_by(., release_year) %>% 
  filter(., !release_year == 2021) %>% 
  summarise(., total_num_user_reviews = sum(num_user_reviews)) %>% 
  ggplot(., aes(x = release_year, y = total_num_user_reviews))+
  geom_col() +
  labs(title = "Number of Videogame Review Scores given by Users\n 1995-2020",
       x= "User Review Scores", y = "Number of User Reviews") +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))
user.reviews.freq


critic.reviews.freq <- game.reviews.df %>%
  group_by(., release_year) %>% 
  filter(., !release_year == 2021) %>% 
  summarise(., total_num_critic_reviews = sum(num_critic_reviews)) %>% 
  ggplot(., aes(x = release_year, y = total_num_critic_reviews))+
  geom_col()+
  labs(title = "Distribution of Videogame Review Scores given by Critics each year\n
       1995-2020",
       x= "Critic Review Scores", y = "Number of Critic Reviews") +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))

critic.reviews.freq


# Over time, the number of user reviews has risen sharply (to be expected)
# User reviews also dwarf critic reviews (400K user reviews in 2020 compared to 20K critic reviews)

```

## 2.4 Inspection: 
  - Which platforms have the highest reviews, on average according to 1) critics, 2) users?
  - Which platform has the most games ranked in the top 100? Bottom 100?

```{r}

# Platform Reviews ####
platform.reviews <- game.reviews.df %>% 
  group_by(., platform) %>% 
  summarise(., avg.annual.critic.score = mean(critic_score_scaled),
            avg.annual.user.score = mean(user_score)) %>% 
  pivot_longer(., cols = !platform, names_to = "review_type", values_to = "review_score")

  # Platform with highest reviews according to Critics
platform.critic.reviews.chart <- platform.reviews %>%
  filter(., review_type == "avg.annual.critic.score") %>% 
  ggplot(., aes(x = reorder(platform, review_score), y =review_score, fill = platform))+
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(review_score,2), hjust = 1.2, size = 1, face = "bold")) +
  labs(title = "Average Critic Review Scores for each Platform",
    subtitle = "Scores out of 10",
    x= "Platform", 
    y = "Average Critic Review Score") +
  scale_fill_discrete() +
  theme(axis.text.x = element_text(size = 10), axis.title.x = element_text(size = 12),
        axis.text.y = element_text(size = 10, face = "bold"), axis.title.y = element_text(size = 12),
        plot.title = element_text(size = 14, face = "bold", hjust = 0.5,),
        plot.subtitle = element_text(hjust = 0.5),
        legend.position = "none") +
  coord_flip()
  # geom_hline(yintercept = mean(game.reviews.df$critic_score_scaled, 
  #                               linetype = "dashed", color = "purple", size = 1.1))

platform.critic.reviews.chart

  # User Reviews
platform.user.reviews.chart <- platform.reviews %>%
  filter(., review_type == "avg.annual.user.score") %>% 
  ggplot(., aes(x = reorder(platform, review_score), y = review_score, fill = platform))+
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(review_score,2), hjust = 1.2, size = 0.7, face = "bold")) +
  labs(
    title = "Average User Review Scores for each Platform",
    subtitle = "Scores out of 10",
    x= "Platform", 
    y = "Average User Review Score") +
  scale_fill_discrete() +
  theme(axis.text.x = element_text(size = 10), axis.title.x = element_text(size = 12),
        axis.text.y = element_text(size = 10, face = "bold"), axis.title.y = element_text(size = 12),
        plot.title = element_text(size = 14, face = "bold", hjust = 0.5,),
        plot.subtitle = element_text(hjust = 0.5),
        legend.position = "none")+
  coord_flip()
  #geom_hline(yintercept = mean(game.reviews.df$user_score, 
                               #linetype = "dashed", color = "purple", size = 1.1))

platform.user.reviews.chart


```

```{r}

# Developer Reviews ####
developer.reviews <- game.reviews.df %>% 
  group_by(., developer) %>% 
  summarise(., 
            avg.annual.critic.score = mean(critic_score_scaled),
            avg.annual.user.score = mean(user_score),
            num.games.developed = n()) %>% 
  pivot_longer(., cols = !c(developer, num.games.developed), names_to = "review_type", values_to = "review_score")

  # Developers with highest reviews according to Critics (>5 games)
developer.critic.reviews.chart <- developer.reviews %>%
  filter(., review_type == "avg.annual.critic.score", num.games.developed > 5) %>% 
  slice_max(., n= 10, order_by = review_score) %>% 
  ggplot(., aes(x = reorder(developer, review_score), y =review_score, fill = developer))+
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(review_score,2), hjust = 1.2, size = 0.6, face = "bold")) +
  labs(
    title = "Highest Rated Game Developers according to Critics",
    subtitle = "Minimum 5 games developed",
    x= "Developer", 
    y = "Average Critic Score out of 10") +
  scale_fill_discrete() +
  theme(axis.text.x = element_text(size = 11), axis.title.x = element_text(size = 12),
        axis.text.y = element_text(size = 10, face = "bold"), axis.title.y = element_text(size = 12),
        plot.title = element_text(size = 14, face = "bold", hjust = 0.5,),
        plot.subtitle = element_text(hjust = 0.5),
        legend.position = "none") +
  coord_flip()
  # geom_hline(yintercept = mean(game.reviews.df$critic_score_scaled, 
  #                               linetype = "dashed", color = "purple", size = 1.1))


  # User Reviews
developer.user.reviews.chart <- developer.reviews %>%
  filter(., review_type == "avg.annual.user.score", num.games.developed > 5) %>% 
  slice_max(., n= 10, order_by = review_score) %>% 
  ggplot(., aes(x = reorder(developer, review_score), y = review_score, fill = developer)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(review_score,2), hjust = 1.2, size = 0.6, face = "bold")) +
  labs(
    title = "Highest Rated Game Developers according to Users",
    subtitle = " Minimum 5 games developed",
    x= "Developer", 
    y = "Average User Score out of 10") +
  scale_fill_discrete() +
  theme(axis.text.x = element_text(size = 11), axis.title.x = element_text(size = 12),
        axis.text.y = element_text(size = 10, face = "bold"), axis.title.y = element_text(size = 12),
        plot.title = element_text(size = 14, face = "bold", hjust = 0.5,),
        plot.subtitle = element_text(hjust = 0.5),
        legend.position = "none")+
  coord_flip() 
  
  #geom_hline(yintercept = mean(game.reviews.df$user_score, 
                               #linetype = "dashed", color = "purple", size = 1.1))

developer.critic.reviews.chart
developer.user.reviews.chart


```


## 2.5 Inspection: Genres
- top reviewed genres, users vs critics
- chart showing freq of genre vs score

```{r}

# Create a dataframe showing the number of genres that each title is associated with
count.genres.per.game <- genres.long %>% 
  filter(., !genre.value == " ") %>% 
  group_by(., platform, title) %>%
  summarize(num_genres = n())

count.genres.per.game



# Table showing genre reviews
all.genre.reviews <- genres.long %>% 
  rename(., "Genre" = genre.value) %>% 
  group_by(., Genre) %>% 
  summarise(., 
            'Frequency' = n(),
            'Average Critic Score' = round(mean(critic_score_scaled), 2),
            'Average User Score' = round(mean(user_score),2))

# Chart showing frquency against average review
ggplot(all.genre.reviews, aes(x = `Frequency`, y = `Average Critic Score`))+
  geom_point()


# Table showing Genres with the highest Critic Reviews (minimum = 1000 games)
top.critic.genres.1000 <- all.genre.reviews %>% 
  filter(., Frequency > 1000) %>% 
  slice_max(., n=5, order_by =`Average Critic Score`) %>% 
  select(., !('Average User Score'))

# Table showing Genres with the highest Critic Reviews (minimum = 100 games)
top.critic.genres.100 <-all.genre.reviews %>% 
  filter(., Frequency > 1000) %>% 
  slice_max(., n=5, order_by =`Average Critic Score`) %>% 
  select(., !('Average User Score')) 

# Table showing Genres with the highest User Reviews (minimum = 1000 games)
top.user.genres.1000 <- all.genre.reviews %>% 
  filter(., Frequency > 1000) %>% 
  slice_min(., n=5, order_by =`Average User Score`) %>% 
  select(., !('Average Critic Score')) 

# Table showing Genres with the highest User Reviews (minimum = 100 games)
top.user.genres.100 <-all.genre.reviews %>% 
  filter(., Frequency > 1000) %>% 
  slice_min(., n=5, order_by =`Average User Score`) %>% 
  select(., !('Average Critic Score')) 

# Format tables

top.critic.genres.1000 <- formattable(top.critic.genres.1000,
                                      align = c("l", "c", "c"))

top.critic.genres.100 <- formattable(top.critic.genres.100,
                                      align = c("l", "c", "c"))

top.user.genres.1000 <- formattable(top.user.genres.1000,
                                      align = c("l", "c", "c"))

top.user.genres.100 <- formattable(top.user.genres.100,
                                   align = c("l", "c", "c"))



```


##2.6 Inspection: Hall of Fame 
- top 5 platforms and top 5 games according to users and critics

```{r}
  # Which were the Top 10 highest reviewed games by Critics?
top10.critic.games  <- game.reviews.df %>%
  slice_min(., n=10, order_by = rank) %>% 
  select(., c(title, critic_score_scaled, platform, release_year))

top10.critic.games.tbl <- formattable(top10.critic.games,
                                   align = c("l", "c", "c"))
top10.critic.games.tbl

  # Which were the Top 10 highest reviewed games by Users (min = 1000 reviews)?
top10.user.games  <- game.reviews.df %>%
  filter(., num_user_reviews > 1000) %>% 
  slice_max(., n=10, order_by = user_score) %>% 
  select(., c(title, user_score, platform, release_year))

top10.user.games.tbl <- formattable(top10.user.games,
                                   align = c("l", "c", "c"))

top10.user.games.tbl
top10.critic.games.tbl


```
